detect patches

### 2024-01-25 scripts/classify_patches.Rmd
fn to save patches rasters

```{r}
p <- '../results/suitability_maps'
r <- rast(file.path(p, 'Acamptopappus_sphaerocephalus1k-2024-01-10_08:57:46.tif'))
msk <- ifel(r < 0.80, NA, r)
m <- mask(r, msk)
m1 <- aggregate(m, 2, fun = 'mean')

@@

patcheR <- functions(x){

  # read in file
  r <- rast(f)
  taxon <- gsub('1k.*$', '', basename(f))
  pout <- '../results/patches'

  # mask raster values < 0.8
  r <- mask(r, ifel(r < 0.80, NA, r))
  r1 <- aggregate(r, 2, fun = 'mean')

  # burn away rivers, and burn patches to hydrologic units ~ populations
  r1 <- mask(r1, hydro, inverse = TRUE)

  # detect patches
  pat <- terra::patches(r1,  directions = 4, allowgaps = FALSE)

  # remove patches < 5 acres
  sizes <- zonal(cellSize(pat, unit="ha"), pat, sum, as.raster=TRUE)
  size_mask <- ifel(sizes < 2.0235, NA, sizes)
  r1 <- mask(r1, size_mask)

  # resample the patches to the original raster resolution
  resample(r1, r, threads = 16, method = 'near',
           filename = file.path(pout, paste0(taxon, '.tif')))

@@

>>>>

```{r}
p <- '../results/suitability_maps'
r <- rast(file.path(p, 'Acamptopappus_sphaerocephalus1k-2024-01-10_08:57:46.tif'))
msk <- ifel(r < 0.80, NA, r)
m <- mask(r, msk)

m <- mask(m, crop(hydr_bound, m), inverse = TRUE)
plot(m)
m1 <- aggregate(m, 2, fun = 'mean')

@@

gd <- '/media/steppe/hdd/Geospatial_data'
hydr_bound <- rast( file.path(gd, 'WBD_rast', 'hu10R-line.tif') )
hydr_ftr <-  rast( file.path(gd, 'river_resistance', 'water_ftR.tif') )

patcheR <- function(x){

  # read in file
  r <- rast(x)
  taxon <- gsub('1k.*$', '', basename(x))
  pout <- '../results/patches'

  # mask raster values < 0.8
  r <- terra::mask(r, ifel(r < 0.80, NA, r))

    # burn away rivers, and burn patches to hydrologic units ~ populations
  r <- terra::mask(r, terra::crop(hydr_bound, r), inverse = TRUE)
  r <- terra::mask(r, terra::crop(hydr_ftr, r), inverse = TRUE)

  # detect patches
  r1 <- terra::aggregate(r, 2, fun = 'mean')
  pat <- terra::patches(r1,  directions = 4, allowGaps = FALSE)

  # remove patches < 5 acres
  sizes <- terra::zonal(cellSize(pat, unit="ha"), pat, sum, as.raster=TRUE)
  size_mask <- terra::ifel(sizes < 4.0470, NA, sizes)
  r1 <- mask(r1, size_mask)

  # resample the patches to the original raster resolution
  terra::resample(r1, r, threads = 16, method = 'near',
           filename = file.path(pout, paste0(taxon, '.tif')))

@@

lapply(f[4:335], patcheR)
